Test susieI using whole genome, s40.22 samples

### 2020-10-24 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
qqplot

The genotype data we used is from UKB biobank, randomly selecting `r as.character(n.ori)` samples. We then filtered samples based on relatedness, that ended up with n = `r as.character(n)` samples.  We use SNP genotype data from chr 1 to chr 22 combined from UKB. SNPs are downsampled to 1/10 (randomly), eQTLs (see below for definition of eQTL) were added back. This ends up with p = `p as.charater(p)` SNPs.

@@

```{r dirs}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
outputdir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
susiedir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
simtag <- '20201001-1-3'
source('~/causalTWAS/causal-TWAS/code/gwas.R')
source('~/causalTWAS/causal-TWAS/code/ld.R')
```

```{r causal}
exprgwasf <- paste0(simdatadir, simtag, ".exprgwas.txt.gz")
load(paste0(simdatadir, "simu_", simtag, "-pheno.Rd"))
caulist <- list()
for (chrom in 1:22) {
  load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40000.FBM.Rd"))
  load(paste0(simdatadir, "simu_s40000_GTEXadipose-B", chrom, "-cis-expr.Rd"))
  caulist[[chrom]]<- c(exprres$gnames[phenores$batch[[chrom]]$param$idx.cgene], dat$snp[phenores$batch[[chrom]]$param$idx.cSNP,])
}
cau <- unlist(caulist)
```

@@

```{r gene twas, fig.height=8, fig.width=8}
chrom <- 22
a <- read.table(exprgwasf, header = T)
a$ifcausal <- ifelse(a$MARKER_ID %in% cau, 1, 0)
ax <- pretty(0:max(-log10(a$PVALUE)), n = 30)

par(mfrow=c(3,1))
h1 <- hist(-log10(a$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey"); grid()
h2 <- hist(-log10(a[a$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon");grid()

@@

@@

par(mfrow=c(3,1))
h1 <- hist(-log10(b$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey"); grid()
h2 <- hist(-log10(b[b$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon");grid()

@@

@@

```{r param1, fig.height=6, fig.width=6}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")

>>>>

The genotype data we used is from UKB biobank, randomly selecting `r as.character(n.ori)` samples. We then filtered samples based on relatedness, ethics and other qc metrics, that ended up with n = `r as.character(n)` samples.  We use SNP genotype data from chr 1 to chr 22 combined from UKB. SNPs are downsampled to 1/10 (randomly), eQTLs (see below for definition of eQTL) were added back. This ends up with p = `p as.charater(p)` SNPs.

@@

```{r dirs}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
outputdir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
susiedir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
simtag <- '20201001-1-3'
source('~/causalTWAS/causal-TWAS/code/gwas.R')
source('~/causalTWAS/causal-TWAS/code/ld.R')
source('~/causalTWAS/causal-TWAS/code/qqplot.R')
```

```{r causal}
exprgwasf <- paste0(simdatadir, simtag, ".exprgwas.txt.gz")
load(paste0(simdatadir, "simu_", simtag, "-pheno.Rd"))
caulist <- list()
for (chrom in 1:22) {
  load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40.22.FBM.Rd"))
  load(paste0(simdatadir, "simu_s40000_GTEXadipose-B", chrom, "-cis-expr.Rd"))
  caulist[[chrom]]<- c(exprres$gnames[phenores$batch[[chrom]]$param$idx.cgene], dat$snp[phenores$batch[[chrom]]$param$idx.cSNP,])
}
cau <- unlist(caulist)
```
# Power estimation

We use gene.pve ~ 0.1, snp.pve ~ 0.5.

* For SNPs, we use $\pi_1 = 2.5e-3$, variance for effect size ~ $0.03^2$, power at 5e-8 p value cutoff:
```{r pow func} 
pow <- function(total, n, beta, cutp){
  rec <- rep(0, total)
  for (i in 1:total){
    x <- rnorm(n)
    y <- x * rnorm(1, sd = beta) + rnorm(n, sd = sqrt(2.5))
    lm.s <- lm(y~x)
    pv <- summary(lm.s)$coefficients[2,4]
    rec[i] <- pv
  }
  length(rec[rec < cutp])/length(rec)
}
```

```{r snp}
total <- 1e3
n <- 22542
p1 <- pow(total, n, 0.0276, 5e-8)
print(p1)
```

* For genes, under low power setting, $\pi_1 = 0.05$, variance for effect size ~ $0.025^2$, power at 1e-5 cutoff:

```{r gene1}
p2 <- pow(total, n, 0.025, 1e-5)
print(p2)
```

For genes, under high power setting, $\pi_1 = 0.02$, variance for effect size ~ $0.045^2$, power at 1e-5 cutoff:

```{r gene2}
p3 <- pow(total, n, 0.045, 1e-5)
print(p3)
```

@@

```{r gene twas, fig.height=8, fig.width=8}
chrom <- 1
a <- read.table(exprgwasf, header = T)
a$ifcausal <- ifelse(a$MARKER_ID %in% cau, 1, 0)
ax <- pretty(0:max(-log10(a$PVALUE)), n = 30)

par(mfrow=c(3,1))
h1 <- hist(-log10(a$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey", xlim= c(3,20), ylim =c(0,50)); grid()
h2 <- hist(-log10(a[a$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon", xlim= c(3,20), ylim =c(0,50));grid()

cat("number of genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5,]))

@@

* qq plot for genes
```{r qqgene, fig.height= 5, fig.width=5}
gg_qqplot(a$PVALUE) +
  theme_bw(base_size = 24) +
  theme(
    axis.ticks = element_line(size = 0.5),
    panel.grid = element_blank()
    # panel.grid = element_line(size = 0.5, color = "grey80")
  )
```

@@

par(mfrow=c(3,1))
h1 <- hist(-log10(b$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey", xlim= c(3,20), ylim =c(0,100)); grid()
h2 <- hist(-log10(b[b$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon", xlim= c(3,20), ylim =c(0,100));grid()

cat("number of SNPs < 5e-8: ", nrow(b[b$PVALUE < 5e-8,]))

@@

* qq plot for SNPs
```{r qqsnp, fig.height= 5, fig.width=5}
gg_qqplot(b$PVALUE) +
  theme_bw(base_size = 24) +
  theme(
    axis.ticks = element_line(size = 0.5),
    panel.grid = element_blank()
    # panel.grid = element_line(size = 0.5, color = "grey80")
  )
```

@@

```{r param1, fig.height=6, fig.width=6}
outputdir <- "~/causalTWAS/simulations/simulation_susieI_20201001/L=2/"
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")
